---
title: "S5_ITS_beta_div"
author: "Stephen Formel"
date: "3/28/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description and notes

Run PERMANOVA and make PCA, PCoA, and NMDS of ITS based on Aitchison

## Load Libraries
```{r, echo = FALSE, message = FALSE}
library(readxl)
library(plyr)
library(tidyverse)
library(phyloseq)
library(data.table)
library(reshape2)
library(car)
library(cowplot)
library(compositions)
library(vegan)
library(ggcorrplot)
library(ggforce)
library(gganimate)
```

## Set up aesthetic for plotting

```{r}

#color palette
cPAL <- c("#F0E442", "#0072B2") #yellow and blue
cPAL <- c("#D55E00", "#0072B2")

SF1_aes <- aes(x = PC1,
               y = PC2,
               shape = orig_soil,
               fill = oil_added)

my.scale_aes <- list(geom_point(size = 2),
                     #geom_polygon(alpha = 0.5),
                     scale_shape_manual(values = c(21,24)),
                     scale_fill_manual(values = c(cPAL, "black", "gray")),
                     theme_bw(),
                     guides(fill = guide_legend(override.aes=list(shape=21, size = 4)),
                            shape = guide_legend(override.aes=list(size = 4))),
                     #labs(shape = "Soil",
                      #    fill = "Oil"),
                     theme(plot.title = element_text(hjust = 0.5)))


```

## Import ITS phyloseq object and clean

```{r}

#read in data
source("../ITS/import_and_clean_S5_ITS.R")

S5.exp <- prune_samples(sample_data(S5)$special_char=="Experimental Sample", S5)
S5.all <- S5

S5 <- S5.exp
```

### Plot of Inocula and Soil together

```{r}

#takes about 5 min on SF macbook pro
df.all <- otu_table(S5.all) %>% 
  as.data.frame()

df.meta.all <- sample_data(S5.all) %>% 
  data.frame()

#make generic for downstream analysis
df <- df.all
df.meta <- df.meta.all

clr.df <- clr(df)
pca = prcomp(clr.df, center = FALSE, scale. = FALSE)

PCA.loadings=pca$rotation
PCA.scores = pca$x

#For below analyses
comm.df <- clr.df %>% 
  as.data.frame()

#for PERMANOVA with all chemistry
comm.df.ITS <- comm.df
```

#### Plot PCA - Aitchison
All tissues + Inocula

```{r}


#make variable for inocula

df.meta$orig_soil[str_detect(df.meta$sampleID_stem, pattern = "O$")] <- "Prev-Oiled Inoc."
df.meta$orig_soil[str_detect(df.meta$sampleID_stem, pattern = "U$")] <- "Not Prev-Oiled Inoc"


#get summary data
pca.sum <- as.data.frame(summary(pca)$importance)
df.PCA <- cbind(PCA.scores,df.meta)

ggplot(data = df.PCA,
       aes(x = PC1,
           y = PC2,
           shape = orig_soil,
           fill = tissue)) +
  geom_point(size = 2) +
  scale_shape_manual(values = c(21,24)) +
  theme_bw() +
  guides(fill = guide_legend(override.aes=list(shape=21, size = 4)),
         shape = guide_legend(override.aes=list(size = 4))) +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title = "PCA on ITS Metagenome - Aitchison",
       caption = "Ellipses are visual guides and have no statistical meaning",
       x = paste("PC1 (", round(pca.sum$PC1[2]*100, digits = 1), "%)"),
       y = paste("PC2 (", round(pca.sum$PC2[2]*100, digits = 1), "%)"),
       shape = "Inoculum") +
  ggforce::geom_mark_ellipse(aes(color = tissue), 
                             alpha = 0.1)

ggsave("S5_ITS_Inocula_PCA.png", width = 8, height = 6)
```

### Plot and PERMANOVA (adonis) of ITS - Aitchison

By Tissue through time
```{r}

#takes about 5 min on SF macbook pro
df.all <- otu_table(S5) %>% 
  as.data.frame()

df.meta.all <- sample_data(S5) %>% 
  data.frame()

#make generic for downstream analysis
df <- df.all
df.meta <- df.meta.all

clr.df <- clr(df)
pca = prcomp(clr.df, center = FALSE, scale. = FALSE)

PCA.loadings=pca$rotation
PCA.scores = pca$x

#For below analyses
comm.df <- clr.df %>% 
  as.data.frame()

#for PERMANOVA with all chemistry
comm.df.ITS <- comm.df
```

#### Plot PCA - Aitchison
All tissues

```{r}

#get summary data
pca.sum <- as.data.frame(summary(pca)$importance)
df.PCA <- cbind(PCA.scores,df.meta)

ggplot(data = df.PCA,
       aes(x = PC1,
           y = PC2,
           shape = orig_soil,
           fill = tissue)) +
  geom_point(size = 2) +
  scale_shape_manual(values = c(21,24)) +
  theme_bw() +
  guides(fill = guide_legend(override.aes=list(shape=21, size = 4)),
         shape = guide_legend(override.aes=list(size = 4))) +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title = "PCA on ITS Metagenome - Aitchison",
       caption = "Ellipses are visual guides and have no statistical meaning",
       x = paste("PC1 (", round(pca.sum$PC1[2]*100, digits = 1), "%)"),
       y = paste("PC2 (", round(pca.sum$PC2[2]*100, digits = 1), "%)"),
       shape = "Inoculum") +
    facet_grid(cols = vars(sampling_period),
               rows = vars(oil_added)) +
  ggforce::geom_mark_ellipse(aes(linetype = orig_soil))

ggsave("S5_ITS_tissue_through_time.png", width = 8, height = 6)
```

## Permanova - Aitchison

Control for sampling period.  This means permutatoins will only happen within a samlping period.  It isn't exactly repeated-measures, but it's in th right d

```{r}

set.seed(1)
mod1 <- adonis(comm.df ~ tissue, 
               data = df.meta, 
               permutations = 9999, 
               method = "euclidean",
               df.meta$sampling_period) 
mod1
```

Only tissue and original soil are significant.

## Soil - Did the inoculation work?

```{r}

S5.soil <- prune_samples(sample_data(S5)$tissue=="Soil", S5)

#takes about 5 min on SF macbook pro
df.soil <- otu_table(S5.soil) %>% 
  as.data.frame()

df.meta.soil <- sample_data(S5.soil) %>% 
  data.frame()

#make generic for downstream analysis
df <- df.soil
df.meta <- df.meta.soil

clr.df <- clr(df)
pca = prcomp(clr.df, center = FALSE, scale. = FALSE)

PCA.loadings=pca$rotation
PCA.scores = pca$x

#For below analyses
comm.df <- clr.df %>% 
  as.data.frame()

#for PERMANOVA with all chemistry
comm.df.ITS <- comm.df

```

#### Plot PCA - Aitchison

Just SOIL
```{r}

#get summary data
pca.sum <- as.data.frame(summary(pca)$importance)
df.PCA <- cbind(PCA.scores,df.meta)

ggplot(data = df.PCA,
       aes(x = PC1,
           y = PC2,
           fill = orig_soil,
           shape = oil_added)) +
  my.scale_aes +
  ggforce::geom_mark_ellipse(aes(linetype = plant_trt),
                             alpha = 0.2) +
  scale_linetype_manual(labels = c("No Plant",
                                   "Planted"),
                        values = c("dashed", "solid")) +
  labs(title = NULL,
       caption = "Ellipses are visual guides and have no statistical meaning",
       shape = "Oil Addition",
       fill = "Inoculum",
       linetype = "Plant Treatment",
       x = paste("PC1 (", round(pca.sum$PC1[2]*100, digits = 1), "%)"),
       y = paste("PC2 (", round(pca.sum$PC2[2]*100, digits = 1), "%)")) +
    facet_grid(cols = vars(sampling_period),
               labeller = labeller(plant_trt = c("Y" = "Planted", "N" = "No Plant"),
                                   sampling_period = c("N16" = "6 Months", "J17" = "1 Year", "N17" = "1.5 Years", "J18" = "2 Years"))) +
  theme(legend.position = "bottom" ,legend.direction = "vertical") +
  scale_color_manual(values = cPAL)

ggsave("S5_ITS_SOIL_through_time.png", width = 8, height = 6)

```
### NMDS

```{r}


A <- comm.df
A <- metaMDS(comm = A, distance = "euclidean")

cbind(df.meta, A$points) %>% 
  ggplot(aes(x = MDS1,
             y = MDS2,
             fill = orig_soil,
           shape = oil_added)) +
  my.scale_aes +
  #ggforce::geom_mark_ellipse(aes(linetype = plant_trt),
  #                           alpha = 0.2) +
  scale_linetype_manual(labels = c("No Plant",
                                   "Planted"),
                        values = c("dashed", "solid")) +
  labs(title = NULL,
       caption = "Ellipses are visual guides and have no statistical meaning",
       shape = "Oil Addition",
       fill = "Inoculum",
       linetype = "Plant Treatment") +
  #   facet_grid(cols = vars(sampling_period),
  #              labeller = labeller(plant_trt = c("Y" = "Planted", "N" = "No Plant"),
  #                                  sampling_period = c("N16" = "6 Months", "J17" = "1 Year", "N17" = "1.5 Years", "J18" = "2 Years"))) +
   theme(legend.position = "bottom" ,legend.direction = "vertical") +
  scale_color_manual(values = cPAL)
```
## Alternative NMDS Plot

The panels I made above were too much and there were some critiques of the PCA, so I've made something that is visually simpler and is an NMDS.  It's hard to see the yellow though, so I'm going to change the colors to something darker.

```{r}


cbind(df.meta, A$points) %>%
  group_by(plant_trt, oil_added, orig_soil, sampling_period) %>% 
  dplyr::summarise(MDS1.mean = mean(MDS1),
            MDS2.mean = mean(MDS2)) %>% 
  ggplot(aes(x = MDS1.mean,
             y = MDS2.mean,
             group = interaction(plant_trt, oil_added, orig_soil))) +
  # geom_link2(aes(alpha = sampling_period),
  #            size = 3) +
  geom_path(aes(color = orig_soil,
                linetype = fct_rev(plant_trt)),
            arrow = arrow(type = "open", 
                          length = unit(2, "mm")),
            size = 0.5,
            show.legend = TRUE) +
  geom_point(data = . %>% 
               filter(sampling_period=="N16"),
             aes(shape = oil_added,
                 fill = orig_soil),
             size = 3) +
  scale_shape_manual(values = c(21,24)) +
  scale_fill_manual(values = c(cPAL, "black", "gray")) +
  scale_color_manual(values = c(cPAL, "black", "gray")) +
  theme_bw() +
  guides(fill = guide_legend(override.aes=list(shape=22, size = 4, linetype = NULL)),
         color = guide_legend(override.aes=list(shape=22, size = 4, linetype = NULL)),
         shape = guide_legend(override.aes=list(shape=c(21,24), size = 4, linetype = NULL))) +
  labs(title = NULL,
       caption = paste0("Stress = ", round(A$stress, 2)),
       shape = "Oil Addition",
       color = "Inoculum",
       fill = "Inoculum",
       linetype = "Plant Presence",
       x = "MDS1",
       y = "MDS1") +
  theme(legend.position = "bottom" ,legend.direction = "vertical",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())


ggsave("S5_ITS_SOIL_through_time_NMDS_vector.png", width = 8, height = 6)
```
## Permanova - Aitchison

Just SOIL
```{r}

set.seed(1)

mod1 <- adonis(comm.df ~ plant_trt*orig_soil*oil_added,
               data = df.meta,
               permutations = 9999,
               method = "euclidean",
               strata = df.meta$sampling_period)

mod1

#check with dbRDA (actually RDA because it's euclidean, but the syntax was more obvious to me in dbRDA than in RDA)

mod1 <- dbrda(comm.df ~ plant_trt + orig_soil + oil_added + Condition(sampling_period), data = df.meta, 
               distance = "euclidean") 

anova(mod1, 
      by = "margin", 
      permutations = 9999)


```

sampling period, plant_trt, orig_soil and the interaction of plant_trt:orig_soil are significant.  But it's still flimsy.  If I take sampling period out, it's still the same, results otherwise.  So I mine as well keep it in.

## dbRDA with oil components + CN

```{r}

mod1 <- dbrda(comm.df ~ pH + conductivity + Hopanes + Total.Naph. + Total.Dibenz. + Total.Phenan. + Total.Chrys. + Condition(sampling_period + table + block), 
              data = df.meta,
              distance = "euclidean") 

mod2 <- dbrda(comm.df ~ 1,
              data = df.meta,
              distance = "euclidean")

ordistep(mod2, scope = formula(mod1))

anova(mod1, 
      by = "margin", 
      permutations = 9999)
```
Significant for Hopanes, pH, table, block.  After controlling for table and block, Hopanes are significant.  Oil addition wasn't significant in the permanova, so that's not very reassuring tha these results mean much.

## Roots - Did the inoculation work?

```{r}

S5.roots <- prune_samples(sample_data(S5)$tissue=="Root", S5)

#takes about 5 min on SF macbook pro
df.roots <- otu_table(S5.roots) %>% 
  as.data.frame()

df.meta.roots <- sample_data(S5.roots) %>% 
  data.frame()

#make generic for downstream analysis
df <- df.roots
df.meta <- df.meta.roots

clr.df <- clr(df)
pca = prcomp(clr.df, center = FALSE, scale. = FALSE)

PCA.loadings=pca$rotation
PCA.scores = pca$x

#For below analyses
comm.df <- clr.df %>% 
  as.data.frame()

#for PERMANOVA with all chemistry
comm.df.ITS <- comm.df

```

#### Plot PCA - Aitchison

Just ROOTS
```{r}

#get summary data
pca.sum <- as.data.frame(summary(pca)$importance)
df.PCA <- cbind(PCA.scores,df.meta)

#remove plant 60, J18 as an outlier

ggplot(data = df.PCA %>% 
         filter(!(plantID=="60" & sampling_period=="J18")),
       aes(x = PC1,
           y = PC2,
           fill = orig_soil,
           shape = oil_added)) +
  my.scale_aes +
  ggforce::geom_mark_ellipse(aes(color = orig_soil),
                             alpha = 0.2) +
  labs(title = NULL,
       caption = "Ellipses are visual guides and have no statistical meaning",
       shape = "Oil Addition",
       fill = "Inoculum",
       color = "Inoculum",
       x = paste("PC1 (", round(pca.sum$PC1[2]*100, digits = 1), "%)"),
       y = paste("PC2 (", round(pca.sum$PC2[2]*100, digits = 1), "%)")) +
    facet_grid(cols = vars(sampling_period),
               labeller = labeller(sampling_period = c("N16" = "6 Months", "J17" = "1 Year", "N17" = "1.5 Years", "J18" = "2 Years"))) +
  theme(legend.position = "bottom" ,legend.direction = "vertical") +
  scale_color_manual(values = cPAL)

ggsave("S5_ITS_ROOT_through_time.png", width = 8, height = 6)

```
## NMDS
```{r}

A <- comm.df[rownames(comm.df)!="S5_R60_J18",]
A <- metaMDS(comm = A, distance = "euclidean")

cbind(df.meta[rownames(df.meta)!="S5_R60_J18",], A$points) %>% 
  ggplot(aes(x = MDS1,
             y = MDS2,
             fill = orig_soil,
           shape = oil_added)) +
  my.scale_aes +
  ggforce::geom_mark_ellipse(aes(linetype = plant_trt),
                             alpha = 0.2) +
  scale_linetype_manual(labels = c("No Plant",
                                   "Planted"),
                        values = c("dashed", "solid")) +
  labs(title = NULL,
       caption = "Ellipses are visual guides and have no statistical meaning",
       shape = "Oil Addition",
       fill = "Inoculum",
       linetype = "Plant Treatment") +
    facet_grid(cols = vars(sampling_period),
               labeller = labeller(sampling_period = c("N16" = "6 Months", "J17" = "1 Year", "N17" = "1.5 Years", "J18" = "2 Years"))) +
  theme(legend.position = "bottom" ,legend.direction = "vertical") +
  scale_color_manual(values = cPAL)
```

## Alternative NMDS Plot

The panels I made above were too much and there were some critiques of the PCA, so I've made something that is visually simpler and is an NMDS.  It's hard to see the yellow though, so I'm going to change the colors to something darker.

```{r}


cbind(df.meta[rownames(df.meta)!="S5_R60_J18",], A$points) %>%
  group_by(oil_added, orig_soil, sampling_period) %>% 
  dplyr::summarise(MDS1.mean = mean(MDS1),
            MDS2.mean = mean(MDS2)) %>% 
  ggplot(aes(x = MDS1.mean,
             y = MDS2.mean,
             group = interaction(oil_added, orig_soil))) +
  # geom_link2(aes(alpha = sampling_period),
  #            size = 3) +
  geom_path(aes(color = orig_soil),
            arrow = arrow(type = "open", 
                          length = unit(2, "mm")),
            size = 0.5,
            show.legend = TRUE) +
  geom_point(data = . %>% 
               filter(sampling_period=="N16"),
             aes(shape = oil_added,
                 fill = orig_soil),
             size = 3) +
  scale_shape_manual(values = c(21,24)) +
  scale_fill_manual(values = c(cPAL, "black", "gray")) +
  scale_color_manual(values = c(cPAL, "black", "gray")) +
  theme_bw() +
  guides(fill = guide_legend(override.aes=list(shape=22, size = 4, linetype = NULL)),
         color = guide_legend(override.aes=list(shape=22, size = 4, linetype = NULL)),
         shape = guide_legend(override.aes=list(shape=c(21,24), size = 4, linetype = NULL))) +
  labs(title = NULL,
       caption = paste0("Stress = ", round(A$stress, 2)),
       shape = "Oil Addition",
       color = "Inoculum",
       fill = "Inoculum",
       x = "MDS1",
       y = "MDS1") +
  theme(legend.position = "bottom" ,legend.direction = "vertical",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

ggsave("S5_ITS_ROOTS_through_time_NMDS_vector.png", width = 8, height = 6)
```
## Permanova - Aitchison

Just ROOTS
```{r}

set.seed(1)
mod1 <- adonis(comm.df ~ orig_soil*oil_added, 
               data = df.meta, 
               permutations = 9999, 
               method = "euclidean",
               strata = df.meta$sampling_period) 
mod1

#check with dbRDA (actually RDA because it's euclidean, but the syntax was more obvious to me in dbRDA than in RDA)

mod1 <- dbrda(comm.df ~ orig_soil + oil_added + Condition(sampling_period), data = df.meta, 
               distance = "euclidean") 

anova(mod1, 
      by = "margin", 
      permutations = 9999)

```

Orig_soil, oil_added, orig_soil:oil_added. dbRDA backs up the main effects idea.

## dbRDA with oil components + CN

```{r}

comm.df.sub <- subset(comm.df, rownames(comm.df)!="S5_R29_N17")
df.meta.sub <- df.meta %>% 
                filter(plantID!=29 | sampling_period!="N17")
                       
mod1 <- dbrda(comm.df.sub ~ percent_N + percent_C + C.N + table + block + + stem_diam + num_nodes + num_stems_live + pH + conductivity + Hopanes + Total.Naph. + Total.Dibenz. + Total.Phenan. + Total.Chrys. + Condition(sampling_period), 
              data = df.meta.sub,
              distance = "euclidean") 

mod2 <- dbrda(comm.df.sub ~ 1,
              data = df.meta.sub,
              distance = "euclidean")

ordistep(mod2, scope = formula(mod1))

anova(mod1, 
      by = "margin", 
      permutations = 9999)
```

Block is the only significant factor.

## Leaves - Did the inoculation work?

Just LEAVES
```{r}

S5.leaf <- prune_samples(sample_data(S5)$tissue=="Leaf", S5)

#takes about 5 min on SF macbook pro
df.leaf <- otu_table(S5.leaf) %>% 
  as.data.frame()

df.meta.leaf <- sample_data(S5.leaf) %>% 
  data.frame()

#make generic for downstream analysis
df <- df.leaf
df.meta <- df.meta.leaf

clr.df <- clr(df)
pca = prcomp(clr.df, center = FALSE, scale. = FALSE)

PCA.loadings=pca$rotation
PCA.scores = pca$x

#For below analyses
comm.df <- clr.df %>% 
  as.data.frame()

#for PERMANOVA with all chemistry
comm.df.ITS <- comm.df

```

#### Plot PCA - Aitchison

Just LEAVES
```{r}

#get summary data
pca.sum <- as.data.frame(summary(pca)$importance)
df.PCA <- cbind(PCA.scores,df.meta)

p <- ggplot(data = df.PCA,
       aes(x = PC1,
           y = PC2,
           fill = orig_soil,
           shape = oil_added)) +
  my.scale_aes +
  ggforce::geom_mark_ellipse(aes(color = orig_soil),
                             alpha = 0.2) +
  labs(title = NULL,
       caption = "Ellipses are visual guides and have no statistical meaning",
       shape = "Oil Addition",
       fill = "Inoculum",
       color = "Inoculum",
       x = paste("PC1 (", round(pca.sum$PC1[2]*100, digits = 1), "%)"),
       y = paste("PC2 (", round(pca.sum$PC2[2]*100, digits = 1), "%)")) +
     facet_grid(cols = vars(sampling_period),
                labeller = labeller(sampling_period = c("N16" = "6 Months", "J17" = "1 Year", "N17" = "1.5 Years", "J18" = "2 Years"))) +
  theme(legend.position = "bottom" ,legend.direction = "vertical") +
  scale_color_manual(values = cPAL)

ggsave("S5_ITS_LEAVES_through_time.png", width = 8, height = 6)


# anim <- p + 
#   transition_states(sampling_period,
#                     transition_length = 2,
#                     state_length = 1)

# anim
```
## NMDS
```{r}

A <- comm.df
A <- metaMDS(comm = A, distance = "euclidean")

cbind(df.meta, A$points) %>% 
  ggplot(aes(x = MDS1,
             y = MDS2,
             fill = orig_soil,
           shape = oil_added)) +
  my.scale_aes +
  ggforce::geom_mark_ellipse(aes(linetype = plant_trt),
                             alpha = 0.2) +
  scale_linetype_manual(labels = c("No Plant",
                                   "Planted"),
                        values = c("dashed", "solid")) +
  labs(title = NULL,
       caption = "Ellipses are visual guides and have no statistical meaning",
       shape = "Oil Addition",
       fill = "Inoculum",
       linetype = "Plant Treatment") +
    facet_grid(cols = vars(sampling_period),
             labeller = labeller(sampling_period = c("N16" = "6 Months", "J17" = "1 Year", "N17" = "1.5 Years", "J18" = "2 Years"))) +
  theme(legend.position = "bottom" ,legend.direction = "vertical") +
  scale_color_manual(values = cPAL)

```
## Alternative NMDS Plot

The panels I made above were too much and there were some critiques of the PCA, so I've made something that is visually simpler and is an NMDS.  It's hard to see the yellow though, so I'm going to change the colors to something darker.

```{r}


cbind(df.meta, A$points) %>%
  group_by(oil_added, orig_soil, sampling_period) %>% 
  dplyr::summarise(MDS1.mean = mean(MDS1),
            MDS2.mean = mean(MDS2)) %>% 
  ggplot(aes(x = MDS1.mean,
             y = MDS2.mean,
             group = interaction(oil_added, orig_soil))) +
  # geom_link2(aes(alpha = sampling_period),
  #            size = 3) +
  geom_path(aes(color = orig_soil),
            arrow = arrow(type = "open", 
                          length = unit(2, "mm")),
            size = 0.5,
            show.legend = TRUE) +
  geom_point(data = . %>% 
               filter(sampling_period=="N16"),
             aes(shape = oil_added,
                 fill = orig_soil),
             size = 3) +
  scale_shape_manual(values = c(21,24)) +
  scale_fill_manual(values = c(cPAL, "black", "gray")) +
  scale_color_manual(values = c(cPAL, "black", "gray")) +
  theme_bw() +
  guides(fill = guide_legend(override.aes=list(shape=22, size = 4, linetype = NULL)),
         color = guide_legend(override.aes=list(shape=22, size = 4, linetype = NULL)),
         shape = guide_legend(override.aes=list(shape=c(21,24), size = 4, linetype = NULL))) +
  labs(title = NULL,
       caption = paste0("Stress = ", round(A$stress, 2)),
       shape = "Oil Addition",
       color = "Inoculum",
       fill = "Inoculum",
       x = "MDS1",
       y = "MDS1") +
  theme(legend.position = "bottom" ,legend.direction = "vertical",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

ggsave("S5_ITS_LEAVES_through_time_NMDS_vector.png", width = 8, height = 6)
```
## Permanova - Aitchison

Just Leaves
```{r}

set.seed(1)
mod1 <- adonis(comm.df ~ orig_soil*oil_added, 
               data = df.meta, 
               permutations = 9999, 
               method = "euclidean",
               strata = df.meta$sampling_period) 
mod1

#check with dbRDA (actually RDA because it's euclidean, but the syntax was more obvious to me in dbRDA than in RDA)

mod1 <- dbrda(comm.df ~ orig_soil + oil_added + Condition(sampling_period), data = df.meta, 
               distance = "euclidean") 

anova(mod1, 
      by = "margin", 
      permutations = 9999)

```


orig_soil, and orig_soil:oil_added are the only levels that are significant. dbRDA supports main effect.

## dbRDA with variables but not treatments

```{r}

comm.df.sub <- subset(comm.df, rownames(comm.df)!="S5_L08_N16")
df.meta.sub <- df.meta %>% 
                filter(plantID!=8 | sampling_period!="N16")
                       
mod1 <- dbrda(comm.df.sub ~ percent_N + percent_C + C.N + table + block +  + stem_ht + stem_diam + num_nodes + num_stems_live + pH + conductivity + Condition(sampling_period), 
              data = df.meta.sub,
              distance = "euclidean") 

mod2 <- dbrda(comm.df.sub ~ 1,
              data = df.meta.sub,
              distance = "euclidean")

ordistep(mod2, scope = formula(mod1))

anova(mod1, 
      by = "margin", 
      permutations = 9999)
```

The only significant thing here is table and block.